##### DEG for getting a logFC for all genes in a sample wise comparitive format #####
library(DESeq2)
library(tidyverse)

# Load data (remove outlier sample AV03)
counts_file = "/ix/djishnu/Akanksha/datasets/bulk_RNA_mouse_liver/ercc1_livers/Ercc1_liver_WD_rawcounts.txt"
counts_data <- read.delim(counts_file, row.names = 1) %>% 
#select(-AV03)  # Remove AV03 column

# Create sample metadata, row names of metadata are taken from column names of raw counts
colData <- data.frame(
  condition = factor(c("WT", "WT", "WT", "WT", "WT", "WT", "WT", "ERCC1_KO", "ERCC1_KO", "ERCC1_KO", "ERCC1_KO", "ERCC1_KO", "ERCC1_KO", "ERCC1_KO")),
  row.names = colnames(counts_data))

# Create DESeq object, with 4 conditions from our dataset
dds <- DESeqDataSetFromMatrix(countData = counts_data, colData = colData, design = ~ condition)

# Filter low gene counts
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]

# Perform the DEG analysis (don't mention reference level at this stage)
dds <- DESeq(dds)
normalized_counts <- counts(dds, normalized = TRUE)
write.csv(normalized_counts, file = "/ix/djishnu/Akanksha/datasets/bulk_RNA_mouse_liver/ercc1_livers/Ercc1_liver_WD_normcounts.csv")
results_deg <- results(dds)

# set the reference and experimental levels
resultsNames(dds)

# Save all DE results to files
res_ERCC1_KO_vs_WT <- results(dds, name = "condition_ERCC1_KO_vs_WT")

# Write results to CSV files
write.csv(res_ERCC1_KO_vs_WT, file = "/ix/djishnu/Akanksha/datasets/bulk_RNA_mouse_liver/ercc1_livers/Ercc1_liver_WD_ERCC1_KO_vs_WT_DEGs.csv")

##### GSEA #####

# order your DEG results to rank the genes
res_LFD_TRF2KO_vs_LFD_CTRL <- res_LFD_TRF2KO_vs_LFD_CTRL[order(-res_LFD_TRF2KO_vs_LFD_CTRL$stat),]
res_WD_CTRL_vs_LFD_CTRL <- res_WD_CTRL_vs_LFD_CTRL[order(-res_WD_CTRL_vs_LFD_CTRL$stat),]
res_WD_KO_vs_WD_CTRL <- res_WD_KO_vs_WD_CTRL[order(-res_WD_KO_vs_WD_CTRL$stat),]

# load libraries for gsea

library(org.Mm.eg.db)
library(clusterProfiler)

# set your deg list 
gene_list_LFD_KO_CTRL <- res_LFD_TRF2KO_vs_LFD_CTRL$stat
names(gene_list_LFD_KO_CTRL) <- rownames(res_LFD_TRF2KO_vs_LFD_CTRL)

gene_list_WD_LFD_CTRL <- res_WD_CTRL_vs_LFD_CTRL$stat
names(gene_list_WD_LFD_CTRL) <- rownames(res_WD_CTRL_vs_LFD_CTRL)

gene_list_WD_KO_CTRL <- res_WD_KO_vs_WD_CTRL$stat
names(gene_list_WD_KO_CTRL) <- rownames(res_WD_KO_vs_WD_CTRL)

# Define the GO term of interest
lipid_biosyn_go_id <- "GO:0008610"
lipid_stor_go_id <- "GO:0019915"
lipid_transport <- "GO:0006869"

# Extract gene symbols associated with the GO term
lipid_biosyn_go_genes <- AnnotationDbi::select(org.Mm.eg.db, keys = lipid_biosyn_go_id, 
                                  keytype = "GOALL", columns = "SYMBOL")
lipid_biosyn_go_gene_set <- unique(lipid_biosyn_go_genes$SYMBOL)
lipid_biosyn_go_gene_set <- lipid_biosyn_go_gene_set[!is.na(lipid_biosyn_go_gene_set)]  # Remove NAs

lipid_stor_go_genes <- AnnotationDbi::select(org.Mm.eg.db, keys = lipid_stor_go_id, 
                                  keytype = "GOALL", columns = "SYMBOL")
lipid_stor_go_gene_set <- unique(lipid_stor_go_genes$SYMBOL)
lipid_stor_go_gene_set <- lipid_stor_go_gene_set[!is.na(lipid_stor_go_gene_set)]  # Remove NAs

lipid_trans_go_genes <- AnnotationDbi::select(org.Mm.eg.db, keys = lipid_transport, 
                                  keytype = "GOALL", columns = "SYMBOL")
lipid_trans_go_gene_set <- unique(lipid_trans_go_genes$SYMBOL)
lipid_trans_go_gene_set <- lipid_trans_go_gene_set[!is.na(lipid_trans_go_gene_set)]  # Remove NAs

```{r}
# List available GO IDs
go_keys <- keys(org.Mm.eg.db, keytype = "GOALL")
head(go_keys)  # Check if your GO ID is present

# Check if the specific GO ID is in the database
"GO:0008610" %in% go_keys
"GO:0019915" %in% go_keys
"GO:0006869" %in% go_keys
```
custom_sets <- list(
  hepatocyte_senescence = c("PLIN2","CIDEC","DGAT1","DGAT2","ACSL1","CIDEA")
)

## (Optional) If you want nicer names on plots:
term2name_df <- data.frame(
  term = names(custom_sets),
  name = c("Hepatocyte Senescence"),
  stringsAsFactors = FALSE
)

## Build TERM2GENE from the lists
term2gene_df <- do.call(rbind, lapply(names(custom_sets), function(tnm) {
  data.frame(term = tnm, gene = unique(custom_sets[[tnm]]), stringsAsFactors = FALSE)
}))

# Convert gene set to TERM2GENE format for each pathway
term2gene_biosyn_df <- data.frame(term = lipid_biosyn_go_id, 
                                  gene = lipid_biosyn_go_gene_set)
term2gene_stor_df <- data.frame(term = lipid_stor_go_id, 
                                gene = lipid_stor_go_gene_set)
term2gene_trans_df <- data.frame(term = lipid_transport, 
                                 gene = lipid_trans_go_gene_set)
# Combine all three pathways into a single TERM2GENE dataframe
term2gene_df <- rbind(term2gene_biosyn_df, term2gene_stor_df, term2gene_trans_df)

# Run GSEA for specific gene set and specific gene list
gsea_result <- GSEA(gene_list_WD_LFD_CTRL, 
                    TERM2GENE = term2gene_df, 
                    minGSSize = 5, 
                    maxGSSize = 700, 
                    pvalueCutoff = 0.5)  # Adjust p-value threshold
# View results
as.data.frame(gsea_result)
```{r}
# Plot GSEA result
library(enrichplot)

# Extract pathway name and p-value
current_pathway <- gsea_result@result$ID[2]
pathway_name <- gsea_result@result$Description[gsea_result@result$ID == current_pathway]
pval <- gsea_result@result$p.adjust[gsea_result@result$ID == current_pathway]
nes <- gsea_result@result$NES[gsea_result@result$ID == current_pathway]

# Generate enhanced GSEA plot
gsea_plot <- gseaplot2(gsea_result, 
          geneSetID = current_pathway, 
          title = paste0("WD CTRL vs LFD CTRL", "\n", pathway_name," - Lipid Storage", " Adj p-value: ", signif(pval, 3)),
          color = "green",
          base_size = 12)  # Adjust font size for better readability
gsea_plot
```

```{r}
png("/ix/djishnu/Akanksha/datasets/bulk_RNA_mouse_liver/gsea_WD_LFD_CTRL_lipid_transport.png", width = 800, height = 600)
print(gsea_plot)
dev.off()
```
